\Large{Adversarial Variational Optimization of Non-Differentiable Simulators}

### 2017-06-19 paper.tex
save

Complex computer simulators are increasingly used across fields of science to
describe generative models tying parameters of an underlying theory to
experimental observations. Inference in this likelihood-free setup is often
intractable, as simulators rarely provide a way to evaluate the likehood
function for a given observation. In this note, we develop a likelihood-free
inference algorithm for fitting a forward non-differentiable generative model to
observed data. We adapt the adversarial training procedure of generative
adversarial networks by replacing the implicit generative network with a
domain-based scientific simulator, and solve the resulting non-differentiable
minimax problem with variational optimization. Effectively, the adversarial
variational optimization procedure results in learning an arbitrarily tight
proposal distribution over simulator parameters, such that the respective
hierarchical model matches the observed data.
\glnote{Mention experimental results.}
\glnote{Add 'so what?' conclusion.}

@@

In many fields of science, computer simulators are used to describe complex
processes that tie parameters of an underlying theory to high dimensional
observations. In most cases, these implicit generative
models~\citep{2016arXiv161003483M} are specified as forward stochastic processes
that generate data. Because it is usually computationally intractable,
simulators often do not provide a way to directly evaluate the likelihood
function for a given observation,  which motivates a class of algorithms for
likelihood-free inference.

@@

The alternating stochastic gradient descent on ${\cal L}_d$ and ${\cal L}_g$ in
GANs inherently assume that the generator $g$ is a differentiable function. In
the setting where we are not interested in learning an implicit model but are
rather interested in the inference of parameters of a fixed
non-differentiable simulator (as outlined in Section~\ref{sec:problem}),
gradients $\nabla_\theta g$ either do not exist or cannot be accessed. As a
result, gradients $\nabla_\theta {\cal L}_g$ cannot be constructed and the
optimization procedure cannot be carried out.

>>>>

Complex computer simulators are increasingly used across fields of science to
describe generative models tying parameters of an underlying theory to
experimental observations. However, inference in this likelihood-free setup is often
intractable, as simulators rarely provide a way to evaluate the likehood
function for a given observation. In this note, we develop a likelihood-free
inference algorithm for fitting a forward non-differentiable generative model to
observed data. We adapt the adversarial training procedure of generative
adversarial networks by replacing the implicit generative network with a
domain-based scientific simulator, and solve the resulting non-differentiable
minimax problem with variational optimization. Effectively, the adversarial
variational optimization procedure results in learning an arbitrarily tight
proposal distribution over simulator parameters, such that the respective
hierarchical model matches the observed data.
\glnote{Mention experimental results.}
\glnote{Add 'so what?' conclusion.}

@@

In many fields of science, computer simulators are used to describe complex
processes that tie parameters of an underlying theory to high dimensional
observations. In most cases, these implicit generative
models~\citep{2016arXiv161003483M} are specified as imperative implementations
of forward stochastic processes that generate data. Because it is usually
computationally intractable, simulators often do not provide a way to directly
evaluate the likelihood function for a given observation,  which motivates a
class of algorithms for likelihood-free inference.

@@

The alternating stochastic gradient descent on ${\cal L}_d$ and ${\cal L}_g$ in
GANs inherently assumes that the generator $g$ is a differentiable function. In
the setting where we are not interested in learning the implicit model itself but are
rather interested in the inference of parameters of a fixed
non-differentiable simulator (as outlined in Section~\ref{sec:problem}),
gradients $\nabla_\theta g$ either do not exist or cannot be accessed. As a
result, gradients $\nabla_\theta {\cal L}_g$ cannot be constructed and the
optimization procedure cannot be carried out.
